Deep Reinforcement Learning for Conservation Decisions

### 2021-06-09 manuscript/manuscript.Rmd
Deleting some sentences in the intro

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or climate change policy models [@dice] are already used to explore scenarios and inform ecosystem management.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.
We believe that the application of RL methods to simulations of ecological processes will open the door to a new paradigm for tackling tough environmental change problems.
We also believe that such a transformation could bring risks and pitfalls comparable and varied as its solutions.

In this paper, we draw on examples from fisheries management and ecological tipping points to illustrate how deep RL techniques can successfully discover optimal solutions to previously solved management scenarios and discover highly effective solutions to unsolved problems.
Our simplest example approaches the known optimal solution for a classic fisheries management problem with no prior knowledge of the underlying model.
Our second example considers ecosystem management under slow environmental degradation: a scenario which is not amenable to classic optimization methods such as dynamic programming.
In this case there is no existing optimal solution to compare against, but an appropriately tuned RL agent can out-perform a sensible rule-of-thumb response.
These examples demonstrate that RL-based approaches are capable but by no means a magic bullet: reasonable solutions require careful design of training environments, choice of RL algorithms, tuning and evaluation, as well as substantial computational power.
Our examples are intentionally simple, aiming to provide a clear template for understanding that could be easily extended to cover more realistic conditions. 
We include an extensive appendix with carefully annotated code which should allow readers to both reproduce and extend this analysis.
We further include implementations of three fully featured simulation modules following current standards for RL benchmark problems as Python modules distributed on PyPi and GitHub.
Each module includes several more challenging variations to the examples described here.
We intend to continue to expand this library of ecological management benchmarks with the hope both of challenging researchers already working in RL to tackle some of these pressing issues and for ecologists and environmental scientists to continue to refine these benchmarks into more realistic ecological simulations.

>>>>

Despite relevance to uncertainty and decision making that could make RL uniquely well suited for ecological and conservation problems, it has so far seen little application in this area.
To date, the problems considered by RL research have largely been drawn from examples in robotic movement and games like Go and Starcraft [@lillicrap; @alphazero; @alphastar]
Complex environmental problems share many similarities to these tasks and games: the need to plan many moves ahead given a large number of possible outcomes, to account for uncertainty and to respond with contingency to the unexpected.
RL agents typically develop their strategies by interacting with simulators, a practice that should not be unsettling to ecologists since learning from simulators is common across ecology.
Rich, processes-based simulations such as the SORTIE model in forest management [@sortie], Ecopath with Ecosim in fisheries management [@ecopath], or climate change policy models [@dice] are already used to explore scenarios and inform ecosystem management.
Decision-theoretic approaches based on optimal control techniques can only find the best strategy in the simplest of ecological models; the so called "curse of dimensionality" makes problems with a large number of states or actions intractable by conventional methods [@Possingham2006; @Marescot2013].
Neural-network-based RL techniques, referred to as *deep RL*, have proven particularly effective in problems involving complex, high-dimensional spaces that have previously proven intractable to classical methods.

In this paper, we draw on examples from fisheries management and ecological tipping points to illustrate how deep RL techniques can successfully discover optimal solutions to previously solved management scenarios and discover highly effective solutions to unsolved problems.
We present multiple examples which demonstrate that RL-based approaches are capable but by no means a magic bullet: reasonable solutions require careful design of training environments, choice of RL algorithms, tuning and evaluation, as well as substantial computational power.
Our examples are intentionally simple, aiming to provide a clear template for understanding that could be easily extended to cover more realistic conditions. 
We include an extensive appendix with carefully annotated code which should allow readers to both reproduce and extend this analysis.
